---
title: "European forecasting hub"
output:
  md_document:
    variant: markdown_github
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source(here::here("template", "report-inputs.R"))
```
### `r format(Sys.Date(), "%d %B %Y")`

We collect, visualise, and combine weekly forecasts of Covid-19 cases and deaths across Europe.

On `r format(forecast_date, "%d %B %Y")`, we collected forecasts from `r team_count` teams. Each team contributed a short-term forecast of new Covid-19 deaths or cases, for at least one of `r location_count` countries. We combined the forecasts using a `r ensemble_method`. See the [visualisation]() or the [raw data]() for detailed forecasts for Europe and each country.

#### Short term situation

Over the next week, forecasters largely expect cases and deaths from Covid-19 to `r country1wk$case$trend[1,1]`, with this trend forecast in `r country1wk$case$trend[1,2]` of `r location_count` countries across Europe. Reports of cases and deaths per 100,000 population are expected to increase in `r pull(filter(country1wk$case$trend, Trend=="increase"), n)` and `r pull(filter(country1wk$death$trend, Trend=="increase"), n)` countries, respectively. 

#### Trends over time {.tabset}

##### Deaths from Covid-19 {.tabset .tabset-pills}

`r paste(country1wk$death$summary_table$Country[1:2])`, and `r country1wk$death$summary_table$Country[3]` are likely to see the highest increases in deaths. 

In `r country1wk$death$summary_table$Country[1]`, deaths are expected to increase  to approximately `r country1wk$death$summary_table[["Forecast weekly per 100,000"]][1]` per 100,000 this week. Over the next four weeks, `r country4wk$death$summary_table$Country[1]` is likely to see the greatest increase  to approximately `r country4wk$death$summary_table[["Forecast weekly per 100,000"]][1]` deaths per 100,000 a week by `r format(country4wk$death$dates[2], "%d %B")`.

###### Region
```{r}
# Summarise by region using values per 100,000
plot(country4wk$death$summary_region_plot) 
```

###### Country

See our [main visualisation](https://covid19forecasthub.eu/visualisation/) for maps and country level plots of forecasts over the next four weeks.

The table below shows forecasts for the 5 countries with highest incidence per 100,000 population. Forecasts were made this week, for the period `r paste(format(country1wk$death$dates[1], "%d %B"), "to", format(country1wk$death$dates[2], "%d %B %Y"))`.

```{r}
knitr::kable(country1wk$death$summary_table[1:5,],
             format.args = list(big.mark = ","))
```

##### Cases of Covid-19 {.tabset .tabset-pills}

`r paste(country1wk$case$summary_table$Country[1:2])`, and `r country1wk$case$summary_table$Country[3]` are likely to see the highest increases in cases. 

In `r country1wk$case$summary_table$Country[1]`, cases are expected to increase  to approximately `r country1wk$case$summary_table[["Forecast weekly per 100,000"]][1]` per 100,000 this week. Over the next four weeks, `r country4wk$case$summary_table$Country[1]` is likely to see the greatest increase to approximately `r country4wk$case$summary_table[["Forecast weekly per 100,000"]][1]` cases per 100,000 a week by `r format(country4wk$case$dates[2], "%d %B")`.

###### Region

```{r}
plot(country4wk$case$summary_region_plot) 
```

###### Country

See our [main visualisation](https://covid19forecasthub.eu/visualisation/) for maps and country level plots of forecasts over the next four weeks.

The table below shows forecasts for the 5 countries with highest incidence per 100,000 population. Forecasts were made this week, for the period `r paste(format(country1wk$death$dates[1], "%d %B"), "to", format(country1wk$death$dates[2], "%d %B %Y"))`.

```{r}
knitr::kable(country1wk$case$summary_table[1:5,],
             format.args = list(big.mark = ","))
```
