---
output:
  pdf_document:
    extra_dependencies: ["booktabs", "colortbl"]
title: Ensembles
---

```{r setup, include = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(here)
library(kableExtra)
library(forcats)
library(tidyr)

table <- read_csv(here::here("ensembles", "evaluation", "weekly-summary",
			     "evaluation-2021-10-18.csv")) %>%
  filter(!(model == c("EuroCOVIDhub-baseline"))) %>%
  filter(is.finite(rel_wis)) %>%
  filter(weeks_included == "All") %>%
  mutate(variable = sub("inc ", "", target_variable),
         variable = stringi::stri_trans_totitle(variable),
         variable = paste0(variable, "s"))

main_ensembles <- table %>%
  filter(model %in% c("EuroCOVIDhub-mean",
                      "EuroCOVIDhub-median",
                      "EuroCOVIDhub-All_relative_skill_weighted_mean",
                      "EuroCOVIDhub-All_relative_skill_weighted_median")) %>%
  mutate(model = sub("EuroCOVIDhub-All_relative_skill_weighted_", "Weighted ", model),
         model = sub("EuroCOVIDhub-", "Unweighted ", model))
```

# Main ensembles

Main ensembles considered are: unweighted mean/median; mean/median weighted by inverse scaled relative WIS.

```{r wis_plot, fig.width = 10, height = 3, echo = FALSE, fig.cap = "Main ensembles, using all component models and considering all of the history. Each point is one country."}
plot_table <- main_ensembles %>%
	filter(location != "Overall", horizon == 2,
	       target_variable != "inc hosp") %>%
	mutate(model = factor(model),
	       model = forcats::fct_rev(model))
p <- ggplot(plot_table, aes(x = model, y = rel_wis)) +
  geom_boxplot(aes(colour = model)) +
  geom_point(aes(colour = model), alpha = 0.5) +
  scale_colour_brewer(palette = "Set1") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  xlab("Ensemble") +
  ylab("Relative WIS") +
  theme_minimal() +
  theme(legend.position = "none") +
  facet_wrap(~ variable, scales = "free") +
  coord_flip()
suppressWarnings(print(p))
```

```{r wis_table, echo = FALSE}
main_ensembles %>%
  filter(location == "Overall") %>%
  mutate(horizon = paste(horizon, if_else(horizon == 1, "week", "weeks"))) %>%
  select(model, Target = variable, Horizon = horizon, rel_wis) %>%
  pivot_wider(names_from = "model", values_from = "rel_wis") %>%
  arrange(Target) %>%
  select(-Target) %>%
  kbl(booktabs = TRUE, caption = "Predictive performance of main ensembles, as measured by the scaled relative WIS.") %>%
  kable_styling(latex_options = "striped") %>%
  group_rows("Cases", 1, 4) %>%
  group_rows("Deaths", 5, 8)
```

# Variations on main ensembles

We further considered: a _cutoff_ (i.e., creating only ensembles from models that had relative scaled WIS < 1 across all forecast horizons for a given target/country); and _limited history_ (i.e., only considering the last 5/10 weeks of scores for the weighted ensemble, as opposed to the whole history). Each point in the countries below represent one country/method/target triplet at the 2-week horizon, with only one aspect of the ensemble varying between the top and bottom rows, connected by lines.

```{r plot_aspects, echo = FALSE, fig.height = 3, fig.width = 10, fig.cap = "Variations on the main ensembles; each point is a country / method pair. Same countries/methods are connected by a line."}
split <- table %>%
  mutate(cutoff =
	   if_else(grepl("cutoff", model), "Cutoff rel. WIS < 1", "No cutoff"),
         average =
	   if_else(grepl("mean", model), "Mean", "Median"),
         weighted =
	   if_else(grepl("weighted", model), "Weighted", "Unweighted"),
         history =
	   if_else(grepl("[0-9]+", model),
		   paste(suppressWarnings(parse_number(gsub("-", "", model))),
			 "weeks history"),
		   "All history"))
reference <- c(cutoff = "No cutoff",
               average = "Mean",
               weighted = "Unweighted",
               history = "All history")
plot_base <- split %>%
	filter(location != "Overall", horizon == 2,
	       target_variable != "inc hosp") %>%
	mutate(model = factor(model),
	       model = forcats::fct_rev(model))
tests <- c("cutoff", "average", "weighted", "history")
differences <- list()
for (test in tests) {
	plot_table <- plot_base %>%
		filter(location != "Overall") %>%
		select_at(c("variable", "horizon", "location_name",
		            setdiff(tests, test), test, "rel_wis")) %>%
		unite("matches", 2:(ncol(.)-2)) %>%
		pivot_wider(names_from = all_of(test), values_from = "rel_wis") %>%
		pivot_longer(3:ncol(.)) %>%
		group_by(variable, matches) %>%
		mutate(all_there = all(!is.na(value))) %>%
		ungroup() %>%
		filter(all_there) %>%
		select(-all_there)
 	p <- ggplot(plot_table, aes(x = name, y = value)) +
		geom_line(aes(group = matches), alpha = 0.25) +
		geom_point(aes(colour = name)) +
		scale_colour_brewer(palette = "Set1") +
		geom_hline(yintercept = 1, linetype = "dashed") +
		xlab("") +
		ylab("Relative WIS") +
		theme_minimal() +
		theme(legend.position = "none") +
		facet_wrap(~ variable, scales = "free") +
		coord_flip()
	print(p)
	ptw <- plot_table %>%
		pivot_wider()
    alternative_columns <- setdiff(colnames(ptw), c("variable", "matches", reference[[test]]))
    for (column in alternative_columns) {
      raw_differences <- ptw[[column]] - ptw[[reference[[test]]]]
      raw_differences <- raw_differences[is.finite(raw_differences)]
      differences[[column]] <- tibble(mean = round(mean(raw_differences), 2),
                                      median = round(median(raw_differences), 2),
                                      low_48 = round(quantile(raw_differences, 0.26), 2),
                                      high_48 = round(quantile(raw_differences, 0.74), 2),
                                      low_96 = round(quantile(raw_differences, 0.02), 2),
                                      high_96 = round(quantile(raw_differences, 0.98), 2))
    }
}
```

```{r plot_effects, echo = FALSE, fig.cap = "Aggregate differences between ensemble variants; shown are the central 48% and 96% intervals (thick/thin lines) and means of the distribution of difference in rel. WIS across all methods/countries."}
dtb <- bind_rows(differences, .id = "change")
ggplot(dtb, aes(x = change)) +
  geom_point(aes(y = mean), pch = 2, size = 4) +
  geom_linerange(aes(ymin = low_48, ymax = high_48), lwd = 2) +
  geom_linerange(aes(ymin = low_96, ymax = high_96), lwd = 1) +
  xlab("") + ylab("Change in relative WIS") +
  theme_minimal() +
  coord_flip() +
  geom_hline(yintercept = 0, linetype = "dashed")
```

