---
title: 'The European COVID-19 Forecasting Hub: Participation'
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(gh)
library(purrr)
library(dplyr)
library(lubridate)
library(ggplot2)
library(ggrepel)
library(stringr)
library(yaml)
library(fs)
```

##### As of `r format.Date(Sys.Date(), "%d %B %Y")`


```{r get-data}
# Get model names / dates from github ---------------------------
model_names <- gh(paste0("/repos/covid19-forecast-hub-europe/covid19-",
                         "forecast-hub-europe/contents/data-processed/?recursive=1"))
model_names <- transpose(model_names)
model_names <- model_names[["name"]]
model_files <- map(model_names,
                   ~ gh(paste0("/repos/covid19-forecast-hub-europe/covid19-",
                               "forecast-hub-europe/contents/data-processed/",
                               .x, "?recursive=1")))
model_names_dates <- unlist(model_files)
model_names_dates <- model_names_dates[names(model_names_dates) == "name"]

models <- tibble("file" = model_names_dates) |>
  mutate(model = substr(file, 12, nchar(file)-4),
         date = floor_date(as.Date(substr(file, 1, 10)),
                           unit="week")) |>
  select(-file) |>
  # exclude Hub own models
  filter(!grepl("EuroCOVIDhub", model))

models <- models |>
  tidyr::separate(col = model, 
                  sep = "-", into = c("team_name", "model_name"), 
                  extra = "warn", remove = FALSE) |> 
  mutate(model = as.factor(model)) |>
  arrange(desc(date)) |>
  mutate(model_freq = forcats::fct_inorder(model, ordered = TRUE))

### Count over time ---------------------------------------------------------
# tally unique models
tally_model <- models |>
  group_by(date) |>
  summarise(n = n(),
            type = "models") |>
  ungroup() 

# tally unique teams
tally_team <- models |>
  distinct(date, team_name) |> 
  group_by(date) |>
  summarise(n = n(),
            type = "teams") |>
  ungroup()

tally <- bind_rows(tally_model, tally_team) |>
  # remove dates before hub launched
  filter(date >= as.Date("2021-03-07"))

# Model recency -------------------------------------------------------------
models_latest <- models |>
  group_by(model) |>
  summarise(latest_date = max(date)) |>
  mutate(most_recent = case_when(latest_date > Sys.Date() - 13 ~ "current",
                                 latest_date > Sys.Date() - 36 ~ "last month",
                                 latest_date > Sys.Date() - 120 ~ "older",
                                 TRUE ~ "older"),
         most_recent = factor(most_recent, c("current", "last month", "older")))

```


We report on teams' involvement in the European COVID-19 Forecast Hub between 7 March 2021 and `r format.Date(max(models$date), "%d %B %Y")`. We exclude two models created by the Hub each week.

A total `r n_distinct(models$team_name)` independent teams have participated in the Hub over time, contributing `r n_distinct(models$model)` unique models. At most, `r max(tally_team$n)` teams were contributing one or more models to the Hub in any one week. In the last month, we collected forecasts from `r n_distinct(filter(models_latest, most_recent != "older"))` unique models. 

#### Participating teams and models over time

```{r tally-plot}
tally |>
  ggplot(aes(x = date, y = n, col = type, label = type)) +
  geom_point() +
  geom_line() +
  labs(x = NULL, y = NULL, col = "Contributing",
       subtitle = "European COVID-19 Forecast Hub participation",
       caption = "Excludes Hub-generated models") +
  scale_color_brewer(type = "qual", palette = 2) +
  scale_x_date(date_breaks = "2 months",
               labels = scales::label_date_short()) +
  scale_y_continuous(limits = c(0, 40)) +
  theme_bw() +
  theme(legend.position = "bottom")
```

##### Individual teams' participation

```{r names-by-recency, fig.height=9, fig.width=7}
# Plot 
models |>
  left_join(select(models_latest, -latest_date)) |>
  ggplot(aes(x = date, y = model_freq)) +
  geom_vline(aes(xintercept = floor_date(Sys.Date(), week_start = 2, unit = "week")), lty = 3) +
  geom_line(aes(col = most_recent)) +
  labs(x = NULL, y = NULL, col = "Last submission") +
  scale_colour_viridis_d(direction = 1) +
  scale_x_date(date_labels = "%b '%y", limits = c(as.Date("2021-03-08"), NA)) +
  scale_y_discrete(limits = rev) +
  theme_bw() +
  theme(legend.position = "bottom")
```

### Complete model information

```{r table-metadata}
github_repo <- "epiforecasts/covid19-forecast-hub-europe"
team_df <-
  gh::gh(
    "/repos/{github_repo}/contents/{path}", 
    github_repo = github_repo, 
    path = "model-metadata"
  )  |>
  map_chr("download_url") |> 
  set_names() |> 
  map(read_yaml) |> 
  imap_dfr(~ list(
    link = .y,
    authors = toString(purrr::map_chr(.x$model_contributors, "name")),
    model_abbr = .x$model_abbr,
    website_url = .x$website_url,
    team_name = .x$team_name,
    methods = .x$methods
  )) |> 
  mutate(
    md_link = glue::glue("[Metadata]({link})"),
    model_abbr = glue::glue("[{model_abbr}]({website_url})")
  ) |> 
  select(authors, model_abbr, team_name, md_link, methods) |> 
  arrange(tolower(model_abbr))

team_df  |> 
  relocate(
    "Model name" = model_abbr,
    "Authors" = authors,
    "Team" = team_name,
    "Methods" = methods,
    "Complete metadata" = md_link
  ) |> 
  knitr::kable(format = "markdown")
```
