---
params: 
  report_date: "2021-03-29"
always_allow_html: true
output:
  html_document:
    theme: yeti
    self_contained: true
    css: https://covid19forecasthub.eu/css/styles.css
title: "European COVID-19 Forecast Hub Ensemble Report"
date: "`r params$report_date`"
---

```{r setup, include=FALSE}
library(purrr)
library(ggplot2)
library(dplyr)
library(DT)
library(knitr)
library(covidHubUtils)
library(lubridate)
library(here)
library(data.table)
library(readr)
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)

include_ranking <- TRUE
include_forecast_plot <- TRUE
```

```{r load-data}
# load forecasts ---------------------------------------------------------------
forecasts <- load_forecasts(source = "local_hub_repo",
                            hub_repo_path = here(),
                            hub = "ECDC") %>%
  filter(model == "EuroCOVIDhub-ensemble")
setDT(forecasts)
# set forecast date to corresponding submision date
forecasts[, forecast_date :=
              ceiling_date(forecast_date, "week", week_start = 2) - 1]
forecasts <- forecasts[forecast_date >= "2021-03-08"]
forecasts <- forecasts[forecast_date <= as.Date(params$report_date)]

setnames(forecasts, old = c("value"), new = c("prediction"))

# load truth data --------------------------------------------------------------
truth <- map_dfr(.x = c("inc case", "inc death"),
                 .f = ~ load_truth(truth_source = "JHU",
                                   target_variable = .x,
                                   hub = "ECDC")) 
setDT(truth)
truth[, model := NULL]
truth <- truth[target_end_date <= as.Date(params$report_date)]
setnames(truth, old = c("value"), 
         new = c("true_value"))

data <- scoringutils::merge_pred_and_obs(forecasts, truth, 
                                         join = "full")

target_variables <- c(Cases = "inc case", Deaths = "inc death")
```

---

# Forecast visualisation {.tabset .tabset-fade}

Forecasts of cases/deaths per week per 100,000. The date of the tab marks the date on which a forecast was made.

```{r forecast-vis, include = FALSE, eval = include_forecast_plot}

locations <- unique(truth$location_name)
forecast_dates <- rev(as.character(unique(data$forecast_date[!is.na(data$forecast_date)])))

out <- NULL
out <- c(out, knit_child(here::here("reports", "ensemble", "template-plot-ensemble.Rmd")))
```

`r paste(if (include_forecast_plot) knit(text = out), collapse = '\n\n')`

# {.unlisted .unnumbered}

---

# Forecast calibration {.tabset .tabset-fade .tabset-dropdown}

Shown below are PIT histograms for the most recent ensemble forecasts. These show the proportion of true values within each predict quantile (width: 0.2).

```{r calibration, echo = FALSE, include = include_ranking}
gap <- 0.2
quantiles <- seq(gap, 1 - gap, by = gap)
scores <- eval_forecasts(data,
                         summarise_by = c("model", "range", "quantile",
                                          "target_variable", "horizon"),
                         pit_plots = TRUE)


even_quantiles <-
  scores[!is.na(quantile) & round(quantile, 3) %in% round(quantiles, 3)]
setkey(even_quantiles, target_variable, horizon, quantile)
pit <- even_quantiles[, list(quantile = c(quantile, 1),
                             pit_bin = diff(c(0, quantile_coverage, 1))),
               by = c("target_variable", "horizon")]

p <- ggplot(pit, aes(x = quantile - gap / 2, y = pit_bin)) +
  geom_col() +
  theme_light() +
  facet_grid(horizon ~ target_variable) +
  xlab("Quantile") + ylab("Proportion") +
  geom_hline(yintercept = gap, linetype = "dashed")

print(p)
```

---
