---
title: Evaluation
author: Sebastian Funk & Kath Sherratt
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document:
    html_preview: false
---

```{r setup, include = FALSE}
library("readr")
library("here")
library("dplyr")
library("tidyr")
library("ggplot2")
library("scales")
library("lubridate")
```

```{r load_data, include = FALSE}
scores <- read_csv(
  here::here("evaluation", "scores.csv"), show_col_types = FALSE
) |>
  filter(forecast_date > today() - years(2))

ensemble_scores <- scores |>
  filter(model == "EuroCOVIDhub-ensemble")
n_models <- scores |>
  group_by(forecast_date) |>
  summarise(models = n_distinct(model))
cov <- ensemble_scores |>
  group_by(forecast_date, target_variable, horizon) |>
  summarise(cov_50 = mean(cov_50),
            cov_95 = mean(cov_95), .groups = "drop")
cov_overall <- ensemble_scores |>
  group_by(target_variable, horizon) |>
  summarise(cov_50 = mean(cov_50),
            cov_95 = mean(cov_95), .groups = "drop")

evaluation_files <- list.files(
  here::here("evaluation", "weekly-summary"),
  full.names = TRUE
)
evaluation_dates <- sub(
  "^.*(202[0-9]-[0-9]{2}-[0-9]{2}).csv$", "\\1", evaluation_files
)
evaluation <- lapply(evaluation_files, read_csv, show_col_types = FALSE)
## convert weeks_included column to character in all
evaluation <- lapply(evaluation, function(x) {
  x$weeks_included <- as.character(x$weeks_included)
  return(x)
})

names(evaluation) <- evaluation_dates
## remove empty data frames
evaluation <- lapply(evaluation, function(x) {
  if (nrow(x) == 0) {
    return(NULL)
  } else {
    return(x)
  }
})

evaluation <- bind_rows(evaluation, .id = "evaluation_date") |>
  mutate(evaluation_date = as.Date(evaluation_date))

ensemble_evaluation <- evaluation |>
  filter(
    model == "EuroCOVIDhub-ensemble",
    location != "Overall",
    weeks_included == 10, ## 10 weeks of history
    evaluation_date > today() - years(2)
  ) |>
  pivot_longer(starts_with("rel_")) |>
  group_by(evaluation_date, horizon, target_variable, name) |>
  summarise(
    central = median(value, na.rm = TRUE),
    lower = quantile(value, 0.05, na.rm = TRUE),
    upper = quantile(value, 0.95, na.rm = TRUE),
    n = n(), 
    .groups = "drop"
  )

criteria_files <- list.files(
  here::here("code", "ensemble", "EuroCOVIDhub", "criteria"),
  full.names = TRUE
)
criteria_dates <- sub(
  "^.*(202[0-9]-[0-9]{2}-[0-9]{2}).csv$", "\\1", criteria_files
)
criteria <- lapply(criteria_files, read_csv, show_col_types = FALSE)
names(criteria) <- criteria_dates
criteria <- bind_rows(criteria, .id = "ensemble_date") |>
  mutate(ensemble_date = as.Date(ensemble_date))
```

# Data files of forecast scores

The files `scores.csv` contains "raw" scores for each data/forecast point based on the predictive quantiles and where the data ended up relative to them. For each combination of date, target variable and location we report the [Weighted Interval Score](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1008618) (`wis`) as well as its components of `sharpness`, `underprediction` and `overprediction`, the absolute error of the median (`aem`), [Bias](https://doi.org/10.1371/journal.pcbi.1006785) (`bias`), coverage at the 50 and 95% levels (`cov_50`/`cov_95`), and the number of quantiles supplied (`n_quantiles`).

The `weekly-summary` folder contains weekly evaluations that enter the [weekly performance reports](https://covid19forecasthub.eu/reports.html).

# Ensemble performance

## Number of models in ensemble

The number of models used in the ensemble (average across all available locations).

```{r nmodels, echo = FALSE}
nmodels <- criteria |>
  group_by(ensemble_date, location, target_variable) |>
  count() |> group_by(ensemble_date, target_variable) |>
  summarise(avg = mean(n), .groups = "drop")

ggplot(nmodels, aes(x = ensemble_date, y = avg)) +
  geom_line() +
  facet_grid( ~ target_variable) +
  theme_bw() +
  scale_x_date(
    "Ensemble date",
    labels = scales::label_date_short()
  ) +
  ylab("Average number of models")
```


## Coverage

Ensemble coverage at 50% and 95% (dashed reference line), showing coverage at each one to four week horizon over time (black line) and average coverage across the four horizons (dotted reference line).

### 50% Coverage
```{r coverage_50, echo = FALSE}
ggplot(cov, aes(x = forecast_date, y = cov_50)) +
  geom_line() +
  facet_grid(target_variable ~ horizon) +
  theme_bw() +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  geom_hline(
    data = cov_overall, aes(yintercept = cov_50), linetype = "dotted"
  ) +
  ylab("50% coverage") +
  scale_x_date(
    "Forecast date",
    labels = scales::label_date_short()
  ) +
  expand_limits(y = c(0, 1))
```

### 95% Coverage
```{r coverage_95, echo = FALSE}
ggplot(cov, aes(x = forecast_date, y = cov_95)) +
  geom_line() +
  facet_grid(target_variable ~ horizon) +
  theme_bw() +
  geom_hline(yintercept = 0.95, linetype = "dashed") +
  geom_hline(data = cov_overall, aes(yintercept = cov_95), linetype = "dotted") +
  ylab("95% coverage") +
  scale_x_date(
    "Forecast date",
    labels = scales::label_date_short()
  ) +
  expand_limits(y = c(0, 1))
```

## Relative performance

Performance is relative to a baseline (a flat constant projection with expanding uncertainty over time). Performance is shown over the preceding 10 weeks.

### Relative WIS (10 weeks)

```{r wis, echo = FALSE}
ensemble_wis <- ensemble_evaluation |>
  filter(name == "rel_wis")
ggplot(ensemble_wis, aes(x = evaluation_date, y = central, ymin = lower, ymax = upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.25) +
  facet_grid(target_variable ~ horizon) +
  theme_bw() +
  scale_x_date(
    "Evaluation date",
    labels = scales::label_date_short()
  ) +
  ylab("Relative WIS") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  coord_cartesian(ylim = c(NA_real_, 2))
```

### Relative AE (10 weeks)

```{r ae, echo = FALSE}
ensemble_ae <- ensemble_evaluation |>
  filter(name == "rel_ae")
ggplot(ensemble_ae, aes(x = evaluation_date, y = central, ymin = lower, ymax = upper)) +
  geom_line() +
  geom_ribbon(alpha = 0.25) +
  facet_grid(target_variable ~ horizon) +
  theme_bw() +
  scale_x_date(
    "Evaluation date",
    labels = scales::label_date_short()
  ) +
  ylab("Relative WIS") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  coord_cartesian(ylim = c(NA_real_, 2))
```
