<!--- 
- Table with coverage by horizon for this week
- Plot of coverage by target and horizon over time
- PIT histograms
--->
# Forecast calibration

The table and plot below show this week's _coverage_ of the ensemble model at the 50% and 95% level, across the 32 countries. This shows the proportion of observations that fall within a given prediction interval. Ideally, a forecast model would achieve 50% coverage of 0.50 (i.e., 50% of observations fall within the 50% prediction interval) and 95% coverage of 0.95 (i.e., 95% of observations fall within the 95% prediction interval). Values of coverage greater than these nominal values indicate that the forecasts are _underconfident_, i.e. prediction intervals tend to be too wide, whereas values of coverage smaller than these nominal values indicate that the ensemble forecasts are _overconfident_, i.e. prediction intervals tend to be too narrow.

## Coverage (this week)

```{r coverage, echo = FALSE, include = include_calibration}
scores <- scoringutils::eval_forecasts(
  data,
  summarise_by = c("model", "range", "quantile",
                  "target_variable", "horizon",
                  "forecast_date"),
  pit_plots = TRUE
)

coverage <- scores %>%
  dplyr::filter(range %in% c(50, 95)) %>%
  select(range, forecast_date, `Target variable` = target_variable,
         `Forecast horizon` = horizon, coverage) %>%
  distinct() %>%
  mutate(range = paste0(range, "% coverage"),
         `Forecast horizon` =
           paste0(`Forecast horizon`, " week",
                  if_else(`Forecast horizon` == 1, "", "s")),
         `Target variable` = recode_factor(`Target variable`,
                                         `inc case` = "Cases",
                                         `inc death` = "Deaths"),
         coverage = round(coverage, 2))

coverage %>%
  tidyr::pivot_wider(names_from = range, values_from = coverage) %>%
  group_by(`Forecast horizon`) %>%
  filter(forecast_date == max(forecast_date)) %>%
  DT::datatable(extensions = c('FixedColumns', 'Buttons'),
                width = "100%",
                options = list(
                  paging = FALSE,
                  info = FALSE,
                  buttons = c('csv', 'excel'),
                  dom = 'Bfrtip',
                  scrollX = TRUE
                ),
                class = 'white-space: nowrap') %>%
  htmltools::tagList()
```

## Coverage (over time) { .tabset .tabset_fade }

The dashed line indicates the nominal 95% level and dotted line the 50% level. If the ensemble was perfectly calibrated, the corresponding coloured lines would coincide with these.

```{r coverage_time, echo = FALSE, include = include_calibration, results='asis'}
for (variable in names(target_variables)) {
  cat("\n\n###", variable, "{.tabset .tabset_fade}\n\n")
  for (horizon in unique(coverage$`Forecast horizon`)) {
    cat("\n\n####", horizon, "\n\n")
    p <- ggplot(coverage %>%
           filter(`Forecast horizon` == horizon,
                  `Target variable` == variable),
           aes(x = forecast_date, y = coverage, colour = range)) +
      geom_line() +
      geom_point() +
      theme_light() +
      scale_colour_brewer("", palette = "Dark2") +
      xlab("Date of the forecast") +
      ylab("Coverage") +
      geom_hline(yintercept = 0.95, linetype = "dashed", colour = "black") +
      geom_hline(yintercept = 0.50, linetype = "dotted", colour = "black") +
      theme(legend.position = "top") +
      ylim(c(0, 1))
    print(p)
  }
}
```

## PIT histograms

The figures below are _PIT histograms_ for the most recent ensemble forecasts. These show the proportion of true values within each predictive quantile (width: 0.2). If the forecasts were perfectly calibrated, observations would fall evenly across these equally-spaced quantiles, i.e. the histograms would be flat.

```{r pit, echo = FALSE, include = include_calibration}
latest_data <- data %>%
  filter(target_end_date == ymd(params$report_date) - 2)
pit_scores <- scoringutils::eval_forecasts(latest_data,
                         summarise_by = c("model", "range", "quantile",
                                          "target_variable", "horizon"),
                         pit_plots = TRUE)
gap <- 0.2
quantiles <- seq(gap, 1 - gap, by = gap)

even_quantiles <- pit_scores %>%
  filter(!is.na(quantile) & round(quantile, 3) %in% round(quantiles, 3))

pit <- even_quantiles %>%
  arrange(target_variable, horizon, quantile) %>%
  group_by(target_variable, horizon) %>%
  summarise(quantile = c(quantile, 1),
            pit_bin = diff(c(0, quantile_coverage, 1)))

p <- ggplot(pit, aes(x = quantile - gap / 2, y = pit_bin)) +
  geom_col() +
  theme_light() +
  facet_grid(horizon ~ target_variable) +
  xlab("Quantile") + ylab("Proportion") +
  geom_hline(yintercept = gap, linetype = "dashed")

print(p)
```
